## Getting Started

To begin, copy all the text on this screen. Then open RStudio, and in the "File" menu select "New File" then "R Markdown". In the box on the left, click "Presentation". Add a title (in the title box) and your name (in the author box), select "PDF (Beamer)" as the Default Output Format, and click Ok. 

A default markdown file will open. Delete all the text in the default file below the header (everything after line 7) and paste in the text you copied to your clipboard. Make sure that "R Markdown" is selected in the file type drop down menu in the bottom right corner of this pane. Save this file in your working directory.

When you Knit the file, select "Knit to PDF (Beamer)" and you'll have slides instead of the usual document. Use a PDF viewer to open these slides...you are making a PDF document, not a Powerpoint!


## Two Hashtags And A Space Identify New Slides

The text following the hashtag becomes the slide title

Everything below the hashtag line becomes the slide content

## Helpful Formatting Tools

- Dashes become bullets on your slide

     - And tabbed dashes become sub-bullets

- All other markdown commands also work in Beamer:

     - Wrap text in *single asterisks* to italicize
     - Wrap text in **double asterisks** to bold
     - Wrap text in ***triple asterisks*** to italicize and bold
     - Wrap text in `brackets` to highlight it
     

## Presentation Tools

>- A dash preceded by an arrow will be a bulleted line that requires "forwarding" to display

>- To add a pause in front of other text (or figure or plot), use three periods (each separated by a space) on a single line:

. . . 

### Three hashtags will create a subheading

All the same rules apply for the content following a subheading

## Including R Code and Output in a Slide

When you want code and the output to appear on a slide, just use a chunk like we have been using in other markdown documents. Note that you can still run the code chunk in this notebook as well (but you do not have to run it for the output to show up on the slide).

```{r}
summary(cars)
```

## Including R Code and Output in a Slide

If you want the output but not the code to appear on a slide, add the `, echo = FALSE` option to the line where you open the chunk. And remember that the `pander` package makes output look much better!

Summary of `speed` variable:

```{r, echo = FALSE}
library(pander)
pander(summary(cars$speed))
```

Summary of `distance` variable:

```{r, echo = FALSE}
pander(summary(cars$dist))
```

## Suppressing Code Is Particularly Helpful With Figures:

```{r, echo = FALSE, warning = FALSE, message = FALSE}
plot(cars)
```

## Including R Code and Output in a Slide

If you want to include code but not have it run, add the `, eval = FALSE` option to the line where you open the chunk:

```{r, eval = FALSE}
library(tidyverse)

plot <- ggplot(cars, aes(x = speed, y = dist))
plot + geom_point()
```

## Beamer Themes

Beamer slides can use many different layouts (called themes) and color options (called color themes). A visual guidebook to all the possible theme and color theme options is available at:

https://hartwork.org/beamer-theme-matrix/

(Note that the above link is clickable on the slide output!)

I use the "Madrid" theme with a custom color theme. But the default color theme is pretty close.

## Beamer Themes

>- To change the theme of your slides, add "theme: Madrid" to the header of this document in line 6 (above the three dashed lines). You can replace `Madrid` with any of the *row options* at the hartwork site. Note that theme names are capitalized.

>- To change the color theme of your slides, add "colortheme: seahorse" to the header of this document. You can replace `seahorse` with any of the *column options* at the hartwork site. Note that color theme names are ***not*** capitalized.