---
title: "Adjusting Color in HTML Slides in R Markdown"
author: "Matt Lawrence"
date: "4/11/2019"
output: 
     ioslides_presentation:
          css: styles.css
---

## Introducing Formatting in Slides

- Colors and other formatting (borders, font size, etc.) can be adjusted with a custom CSS file. 

- The `styles.css` file on Canvas makes some adjustments that produce slides similar in format to the slides used in class. ***Put the css file in the same folder as the Rmd file you are using to create your slides!***

- Note that if you are using the custom CSS file, the output section in the Markdown file header is slightly different. It should look like this:

```
output:
     ioslides_presentation:
          css: styles.css
```

## Introducing Formatting in Slides

- The colors in the `styles.css` file can be changed to any of the color options available on [***this page***.](https://www.w3schools.com/cssref/css_colors.asp) For example, to change the color of your title text, change "mediumblue" in line five of the css file to one of the options in the "Color Name" column in the linked page.

     - (Note that you can click on the link above in your slide.)

- The remaining slides in this document are the same as those from the original ioslides.Rmd file. But in this file you can see how they change as you adjust the colors.

## Knit To Slides!

To see your slides, it's best to open the file in a browser (preferably Google Chrome).

When the file is open in a browser, you can press `f` to expand to a full screen.

You can also press `w` to expand to a wide screen.

## Two Hashtags And A Space Identify New Slides

The text following the hashtag becomes the slide title

Everything below the hashtag line becomes the slide content

## Helpful Formatting Tools

- Dashes become bullets on your slide

     - And tabbed dashes become sub-bullets

- All other markdown commands also work in presentations:

     - Wrap text in *single asterisks* to italicize
     - Wrap text in **double asterisks** to bold
     - Wrap text in ***triple asterisks*** to italicize and bold
     - Wrap text in `brackets` to highlight it

>- A dash preceded by an arrow will be a bulleted line that requires "forwarding" to display

## Including R Code and Output in a Slide

When you want code and the output to appear on a slide, just use a chunk like we have been using in other markdown documents. Note that you can still run the code chunk in this notebook as well (but you do not have to run it for the output to show up on the slide).

```{r}
summary(cars)
```

## Including R Code and Output in a Slide

If you want the output but not the code to appear on a slide, add the `, echo = FALSE` option to the line where you open the chunk. And remember that the `pander` package makes output look much better!

## Using Pander In Slides

Summary of `speed` variable:

```{r, echo = FALSE}
library(pander)
pander(summary(cars$speed))
```

Summary of `distance` variable:

```{r, echo = FALSE}
pander(summary(cars$dist))
```

## Suppressing Code Is Helpful For Figures

```{r, echo = FALSE}
plot(cars)
```

## Adjusting Figure Size

The default size for plots is somewhat small. The adjustments in the first line of the chunk below will usually create a figure that looks better on the slide.

```{r, eval=FALSE, out.width="900px", out.height="500px"}

plot(cars)
```

See how on the next slide...

## Adjusting Figure Size

```{r, echo = FALSE, out.width="900px", out.height="500px"}
plot(cars)
```


## Including R Code and Output {.build}

If you want to include code but not have it run, add the `, eval = FALSE` option to the line where you open the chunk:

```{r, eval = FALSE}
library(tidyverse)

plot <- ggplot(cars, aes(x = speed, y = dist))
plot + geom_point()
```

On the next slide, we will show the plot but not the code...

## Figure Made With ggplot

```{r, echo = FALSE, out.width="900px", out.height="500px", error = FALSE, warning = FALSE, message = FALSE}

library(tidyverse)
plot <- ggplot(cars, aes(x = speed, y = dist))
plot + geom_point()
```


## Printing Slides

Printing works best when you open the html file in Google Chrome. 

In Chrome, save the file as a PDF and make sure the "Background Graphics" button is selected in the "More Settings" menu.
