## Setting Up

Load the `gss_week7.csv` file on Canvas as a data frame called `gss_week7` and load the usual packages:

```{r, include = FALSE, warning = FALSE}
gss_week7 <- read.csv("gss_week7.csv")
library(tidyverse)
library(pander)
```


## Reviewing Significance Tests For Means 

R has a built-in function called `t.test()` that will calculate the test statistic, the p-value, and the confidence interval of a mean. For a test of one mean, we have to fill in the value of mu which is the null hypothesis value. If you leave it out, the default is zero.

To warm up, test whether the mean number of ideal children (`ideal_childs`) significantly differs from 2.42 (a value previous research suggested was the true population mean in a different sample) at the .05 alpha level.

### REPLACE THIS LINE WITH YOUR CODE

Can you reject the null hypothesis that the mean of `ideal_childs` in the full sample differs from 2.42 at the .01 alpha level?

### REPLACE THIS LINE WITH YOUR CODE


## Significance Tests For Proportions

Proportion tests use a different standard error. The R shortcut (`prop.test()`) also uses a different syntax.

For this exercise, we will use the `kidssol` variable. We will want to test whether the proportion of respondents reporting their children's living standards will be "Much better" or "Somewhat better" is significantly different from .50.

Start by looking at the distribution of responses to the `kidssol` question.

### REPLACE THIS LINE WITH YOUR CODE


There are 701 cases without a value for `kidssol`. We need to replace their blank value with NA so R knows how to handle them correctly. We can do this with mutate and na_if.

```{r}
gss_week7 <- gss_week7 %>%
     mutate(kidssol = na_if(kidssol, "")) %>%
     droplevels() # Delete the unusued factor level
```

Let's look at a proportion table of `kidssol`.

### REPLACE THIS LINE WITH YOUR CODE


Now create a binary variable called `kidssol_better` distinguishing the `kiddsol` values of "Much better" and "Somewhat better" from all other values. Give "Much better" and "Somewhat better" values a 1 for the new variable.

### REPLACE THIS LINE WITH YOUR CODE

What is the mean of this new binary variable?

### REPLACE THIS LINE WITH YOUR CODE

The `prop.test()` function requires the number of cases with a 1 and the total number of cases. To find those, wrap a `table()` function in the `addmargins()` function (to get the total):

```{r}
addmargins(table(gss_week7$kidssol_better))
```

In the `prop.test()` code, the first number is the count of observations with a 1 and the second number is the sum:

```{r}
prop.test(306, 565)
```

Can we reject the null hypothesis that the true proportion is .5?

One more exercise: At the .01 level, can you reject the null hypothesis that the true proportion is .62?

### REPLACE THIS LINE WITH YOUR CODE