## Setting Up

We'll use two different files today: `week_8_1.csv` and `week_8_2.csv`. Load them both with the usual packages.


```{r, echo = FALSE, warning = FALSE, message = FALSE}
library(tidyverse)
library(pander)
week_8_1 <- read.csv("week_8_1.csv")
week_8_2 <- read.csv("week_8_2.csv")
```

## Review of Comparing Differences in Means

Let's start by finishing up the other group exercises from the end of Tuesday's class. We had a quick look at testing whether the mean number of political actions differs by sex. Does it?

```{r means exercise 1}
t.test(week_8_1$polactions[week_8_1$sex=="Male"],
       week_8_1$polactions[week_8_1$sex=="Female"])
```

Does the average age at first birth differ by race?

Is the average age of self employed workers significantly different from the average of those who work for someone else?


## Comparing Proportions
We have seen over the last couple weeks that proportions and means have different standard errors. That is still true when we are comparing differences in proportions. Also, when working with differences in proportions, we will use different SEs for confidence intervals and for significance tests.

The SE for a CI of a difference in proportions is very similar to what we used last week as the SE for a CI of a proportion:

\Large{$se = \sqrt{ \frac {\hat{\pi}_1 (1 - \hat{\pi}_1)} {n_1} + \frac {\hat{\pi}_2 (1 - \hat{\pi}_2)} {n_2} }$}

The SE for a difference in proportions uses a pooled estimate of the proportion and the n's for each sample:

\Large{$se = \sqrt{ \hat{\pi}(1-\hat{\pi})  (\frac {1} {n_1} + \frac{1}{n_2})  }$}

The null hypothesis is that proportions are the same in each sample. Formally, this is written as:

\Large{$\pi_2 - \pi_1 = 0$}

The test statistic is calculated as:

\Large{$\frac{\hat{\pi}_1 - {\hat{\pi}_2}}{se}$}


## Example

Now we'll use the `week_8_2.csv` file. Let's look at racial differences in how respondents think people "get ahead". We can check the GSS website for a description of this variable: https://gssdataexplorer.norc.org/variables/585/vshow


Start with a proportion table with `racehisp` in the rows and `getahead` in the columns:

### REPLACE THIS LINE WITH YOUR CODE ###

Let's see if the proportions believing people get ahead through only hard work differ between Hispanic and white respondents at the .05 alpha level

We need a binary variable (!) distinguishing those who answered "hard work" for the `getahead` variable from those who answered "Luck or help" or "Both equally". Call this new variable `hardwork`. And the rest of the process will be easier if we also have a data frame that only has Hispanic and white respondents. Let's call it `gss2`.

### REPLACE THIS LINE WITH YOUR CODE ###

Now let's look at the table using our new data frame:

### REPLACE THIS LINE WITH YOUR CODE ###

Notice that the table still includes rows for "Black" and "Other", the two levels of `racehisp` that we filtered out. Add `droplevels()` to get rid of the unused factor levels:

```{r review table with droplevels}
gss2 <- week_8_2 %>% 
      mutate(hardwork = 
            ifelse(getahead == "Hard work", 1, 0)) %>%
      filter(racehisp == "Hispanic" | 
            racehisp == "White") %>%
      droplevels()

pander(round(prop.table(table(
      gss2$racehisp, gss2$hardwork),1),3))
```

## Testing Differences in Proportions

Now we are ready to test the differences in proportions. We are going to use the `prop.test()` function again so we need the frequencies. Let's save them in a table called `gss2_table`.

```{r proportions exercise frequency table}
gss2_table <- table(gss2$racehisp, gss2$hardwork)
```

And now feed this saved table into the `prop.test()` function:

```{r prop test with frequency table}
prop.test(gss2_table)
```

At the .05 level, are these proportions significantly different from each other?


## Comparing Proportions - Exercise

At the .05 level, test the difference in the "hardwork" proportions between respondents who self-report as lower class and respondents who self-report as middle class.

### REPLACE THIS LINE WITH YOUR CODE ###
