## Setting Up

We'll use the `week_8_2.csv` file today. Load it with the usual packages.

```{r, echo = FALSE, warning = FALSE, message = FALSE}
library(tidyverse)
library(pander)
week_8_2 <- read.csv("week_8_2.csv")
```


## Comparing Proportions

We have seen over the last couple weeks that proportions and means have different standard errors. That is still true when we are comparing differences in proportions. Also, when working with differences in proportions, we will use different SEs for confidence intervals and for significance tests.

The SE for a CI of a difference in proportions is very similar to what we used last week as the SE for a CI of a proportion:

$\Large{se = \sqrt{ \frac {\hat{\pi}_1 (1 - \hat{\pi}_1)} {n_1} + \frac {\hat{\pi}_2 (1 - \hat{\pi}_2)} {n_2} }}$

The SE for difference uses a pooled estimate of the proportion and the n's for each sample:

$\Large{se = \sqrt{ \hat{\pi}(1-\hat{\pi})  (\frac {1} {n_1} + \frac{1}{n_2})  }}$

The null hypothesis is that proportions are the same in each sample. Formally, this is written as:

$\Large{\pi_2 - \pi_1 = 0}$

The test statistic is calculated as:

$\Large{\frac{\hat{\pi}_1 - {\hat{\pi}_2}}{se}}$


## Example

Let's look at racial differences in how respondents think people "get ahead". We can check the GSS website for a description of this variable: https://gssdataexplorer.norc.org/variables/585/vshow

Start with a proportion table with `racehisp` in the rows and `getahead` in the columns:

### REPLACE THIS LINE WITH YOUR CODE


Let's see if the proportions believing people get ahead through only hard work differ between Hispanic and white respondents at the .05 alpha level.

We need a binary variable distinguishing those who answered "hard work" for the `getahead` variable from those who answered "Luck or help" or "Both equally". Call this new variable `hardwork`. And the rest of the process will be easier if we also have a data frame that only has Hispanic and white respondents. Let's call it `gss2`.

### REPLACE THIS LINE WITH YOUR CODE

Now let's look at the proportion table using our new data frame:

### REPLACE THIS LINE WITH YOUR CODE


Notice that the table still includes rows for "Black" and "Other", the two levels of `racehisp` that we filtered out. Add `droplevels()` to get rid of the unused factor levels:

```{r review table with droplevels}
gss2 <- week_8_2 %>% 
      mutate(hardwork = 
            ifelse(getahead == "Hard work", 1, 0)) %>%
      filter(racehisp == "Hispanic" | 
            racehisp == "White") %>%
      droplevels()

pander(round(prop.table(table(
      gss2$racehisp, gss2$hardwork),1),3))
```

## Testing Differences in Proportions

Now we are ready to test the differences in proportions. We are going to use the `prop.test()` function again so we need the frequencies. 

```{r proportions exercise frequency table}
gss2_table <- table(gss2$racehisp, gss2$hardwork)
gss2_table
```

And now feed this saved table into the `prop.test()` function:

```{r prop test with frequency table}
prop.test(gss2_table)
```

At the .05 level, are these proportions significantly different from each other?


## Comparing Proportions - Exercise

At the .05 level, test the difference in the "hardwork" proportions between respondents who self-report as lower class and respondents who self-report as middle class.

### REPLACE THIS LINE WITH YOUR CODE
