# Setting Up

Load the `counties.csv` file on Canvas as a data frame called `counties` and load the usual packages:

```{r load data and packages, echo = FALSE, warning = FALSE, message = FALSE}
counties <- read.csv("/counties.csv")
library(tidyverse)
library(pander)
```


# Summarizing Multiavariate Relationships

As a warm up, use `group_by` to compare the mean teen birthrates for counties with mobility rates higher than the national average (`mobility_hi` = 1) and for counties with mobility rates lower than the national average (`mobility_hi` = 0).

### REPLACE THIS LINE WITH YOUR CODE



On Tuesday, we discussed how we can begin to explain associations between two variables by controlling for (or holding constant) a third variable. To summarize multivariate descriptives, we can extend `group_by` to make a table with three (or more!) variables. All we have to do is put the two x variables in the `group_by` function, listing the control variable first. 

In this example, the state will be the control variable, and we will filter to only include counties from the states of New York, California, Illinois, Connecticut, and New Mexico for simplicity:

```{r create three way table}
three_way_table <- counties %>%
     filter(state == "New York" | state == "California" |
                 state == "Illinois" | state == "Connecticut" |
                 state == "New Mexico") %>%
     group_by(state, mobility_hi) %>%
     summarize(mean_teen_birthrate = mean(teen_birthrate, na.rm=TRUE))

pander(three_way_table)
```

Other functions we have looked at also work when describing three variables. For example, we could do a proportion table. This time we will look at the proportion of counties in each state that have values for the commute variable and the mobility variable above the mean.

```{r regular table with three variables}
counties_subset <- counties %>%
     filter(state == "New York" | state == "California" |
                 state == "Illinois" | state == "Connecticut" |
                 state == "New Mexico") %>%
     droplevels() #droplevels gets rid of the levels for the factor variable (state) that we are not including in the subset

three_way_table_2 <- table(counties_subset$state, counties_subset$commute_hi, counties_subset$mobility_hi)

pander(three_way_table_2)
```

If you want proportions for a table with three variables, it is preferable to combine `prop.table()` with a flat table using the `ftable()` function.

```{r flat table}
round(prop.table(ftable(counties_subset$state, counties_subset$commute_hi, counties_subset$mobility_hi),1),2)
```


# More On Missing Data

In the tables above, we can see that some counties are missing values for the `teen_birthrate` variable. How do we summarize the distribution of a variable if there are missing values? (Use the full counties dataframe here.)

### REPLACE THIS LINE WITH YOUR CODE 


We can see this dataframe includes 350 counties missing values for the teen_birthrate variable. There are two reasons data can be missing:

- *Missing At Random*: no association between probability of a variable having missing values and the values of another variable

- *Missing Not At Random*: there is an association between the probability of a variable having missing values and the values of another variable. This is a bigger problem!

Usually we would want to investigate whether there are differences in other variables between observations that are missing values and observations that are not missing values. In this case, the counties missing values for the `teen_birthrate` variable are also missing values for all the other variables. Let's drop all those counties using filter, and summarize the distribution to confirm they are gone:

```{r drop missing values}
counties <- filter(counties, !is.na(teen_birthrate))
summary(counties$teen_birthrate)
```




### BACK TO SLIDES; WE'LL RETURN TO R SOON ###




# Introducing Probability And Z-Scores

A z-score or standardized value is a value's distance from the mean in standard deviations. It is calculated as: $\Large{z = \frac{x - \mu}{\sigma}}$. In words, the z-score is the difference between the observed value and the sample mean divided by the standard deviation.

After confirming that `teen_birthrate` is approximately normally distributed, let's make a new variable with the standardized values of `teen_birthrate`:

```{r standardize values}
counties <- mutate(counties,
     teen_birthrate_z = (teen_birthrate - mean(teen_birthrate)) /
                         sd(teen_birthrate))
```

Z-scores should be normally distributed with a mean of 0 and a standard deviation of 1. Were we successful?                        

###REPLACE THIS LINE WITH YOUR CODE###


What is the z-score for Addison County, Vermont's birth rate?

###REPLACE THIS LINE WITH YOUR CODE###



When we plot standardized values that are approximately normal, we now know a lot about how many observations fall along different points of the distribution. Make a density plot showing the distribution of the standardized teen birthrates:

###REPLACE THIS LINE WITH YOUR CODE


Each point on the y-axis is the probability of getting that exact value. To find the probability of getting any z-score, use `dnorm()`. Think about this value as the y axis point on the density curve for any specific value on the x axis. For example, the probability that a randomly pulled county would have the same birthrate as Addison County, Vermont is:

```{r dnorm to find probability of z score}
dnorm(-1.75531)
```


Probabilities of specific values are more helpful for descriptives than for inference. Moving forward, what is more helpful is knowing the probability of randomly pulling a value that is greater than or less than an observed value.  In other words, we want to add up the probabilities of pulling any value less than Addison County's value.

We get that summed probability by thinking not of the density but of the *cumulative density*. The cumulative density is also the percentile.

If you have the z-value and want the percentile associated with it, use `pnorm()`. For Addison County:

```{r pnorm for percentile of z score}     
pnorm(-1.75531)
```

So about 4% of counties have a teen birth rate less than the teen birth rate of Addison County.

We can also plot the entire *empirical cumulative density function*:

```{r ecdf}
birthrate_z_plot + geom_step(stat = "ecdf") + 
     labs(y = "cumulative density")
```


Since normal distributions are symmetrical, the probability of getting a value that is more extreme of a negative z-score is the same as the probability of getting a value that is more extreme of that that positive z-score.

What is the probability of another county having a birth rate that is higher than that of Hamilton County, FL?

###REPLACE THIS LINE WITH YOUR CODE###


What proportion of counties have birth rates between those of Addison and Hamilton?

###REPLACE THIS LINE WITH YOUR CODE###



# Exercise

1. Choose a county and find its z-score for the average commute time (variable = `commute`). 

2. What is the probability that a randomly pulled county will have an average commute time greater than the county you chose?

3. What is the probability that a randomly pulled county will have an average commute time less than the county you chose?