## Setting Up

Load the `gss_week7.csv` file on Canvas as a data frame called `gss_week7` and load the usual packages:

```{r, include = FALSE, warning = FALSE}
gss_week7 <- read.csv("~/gss_week7.csv")
library(tidyverse)
library(pander)
```


## Reviewing Significance Tests For Means 

R has a built-in function called `t.test()` that will calculate the test statistic, the p-value, and the confidence interval. For a test of one mean, we have to fill in the value of mu which is the null hypothesis value. If you leave it out, the default is zero.

Can you reject the null hypothesis that the mean of `ideal_childs` in the full sample differs from 2.42 at the .05 alpha level?

### REPLACE THIS LINE WITH YOUR CODE ###

Can you reject the null hypothesis that the mean of `ideal_childs` in the full sample differs from 2.42 at the .01 alpha level?

### REPLACE THIS LINE WITH YOUR CODE ###


## Significance Tests For Proportions

Proportion tests use a different standard error. The R shortcut (`prop.test()`) also uses a different syntax.

For this exercise, we will use the `kidssol` variable. At the .05 alpha level, is the proportion of respondents reporting their children's living standards will be "Much better" or "Somewhat better" different from .50?

Before we use `ifelse()` to create the binary variable, we first have to fix some missing values. To see the problem we need to fix, make a table of the `kidssol` variable:

### REPLACE THIS LINE WITH YOUR CODE ###

There are 701 cases without a value for `kidssol`. We need to replace their blank value with NA so R knows how to handle them correctly:

```{r}
gss_week7$kidssol[gss_week7$kidssol==""] <- NA
```

Now create a binary variable called `kidssol_better` distinguishing the `kiddsol` values of "Much better" and "Somewhat better" from all other values:

### REPLACE THIS LINE WITH YOUR CODE ###

The `prop.test()` function requires the number of cases with a 1 and the total number of cases. To find those, wrap a `table()` function in the `addmargins()` function (to get the row total):

```{r table with sum using addmargins}
addmargins(table(gss_week7$kidssol_better))
```

In the `prop.test()` code, the first number is the count of observations with a 1 and the second number is the sum:

```{r prop test example 1}
prop.test(306, 565)
```

Can we reject the null hypothesis that the true proportion is .5?

One more exercise: At the .01 level, can you reject the null hypothesis that the true proportion is .62?

### REPLACE THIS LINE WITH YOUR CODE ###
