Today we will continue using data from Chetty et al's 2014 paper "Where Is The Land Of Opportunity?". The `commuting_zones.csv` file on Canvas comes from the Opportunity Insights website which can be accessed [`here`](http://www.opportunityinsights.org).

Load the data as a data frame called `cz` and load the tidyverse and pander packages.

```{r load data and packages, include=FALSE}
cz <- read.csv("/users/lawrence/desktop/soan385_data/commuting_zones.csv")
library(tidyverse)
# install.packages("pander")
# remove the hashtag in the previous line if you need to install pander
library(pander)
```

Last week we began thinking about how two variables are associated by finding whether they were positively or negatively correlated and how weak or strong the correlation is. We looked at the variables measuring the proportion of workers in a commuting zone commuting 15 minutes or less, income, racial segregation, labor force participation, and a measure of upward mobility.

When we finished last week, we were looking at the correlation between `commute15min` and `mobility`. How would you describe the correlation between these two variables?

### REPLACE THIS LINE WITH YOUR CODE

And how can you make a scatterplot showing the correlation between these two variables?

### REPLACE THIS LINE WITH YOUR CODE


The scatterplots Ryan Clement showed us on Thursday included a "line of best fit" to help visualize the direction and strength of correlations. We will do much more with these lines at the end of the semester. But for now, it could be good to know how to include them. To do so, add another layer to the plot with the `geom_smooth(method = lm)` function. The smooth function finds a pattern across all the points, and the `method = lm` option says we want the pattern to be based on the "linear model" used in basic regression models.

```{r mobility commute scatter with line of best fit, warning=FALSE}
mobility_commute_scatter <- ggplot(cz, aes(x = commute15min, y = mobility))

mobility_commute_scatter + geom_point() + geom_smooth(method = lm)
```


Let's spend some time exploring a few of the other associations in our data. To start, use `select()` to create a new data frame called `cz_subset` with the following variables: 

- `cz_name` = commuting zone name
- `state` = commuting zone state
- `mobility` = measure of absolute upward mobility
- `gini` = Gini coefficient of income inequality; higher gini values indicate more inequality
- `social_capital_index` = Social capital index
- `frac_foreign_born` = Proportion of residents born in other countries
- `frac_children_single_mothers` = Proportion of children living in single-parent households
- `school_expenditures_per_student` = Average expenditures per student in public schools
- `hs_dropout_rate_adj` = High school dropout rate adjusted for family income; positive values indicate that the hs dropout rate is larger than expected given a commuting zone's median family income, and negative values indicate that the hs dropout rate is smaller than expected given a commuting zone's median family income
- `urban` = binary variable for urban (1) or rural (0) commuting zone
- `hh_income` = median household family income in commuting zone
- `racial_seg` = measure of racial segregation


```{r create subset}
cz_subset <- cz %>%
     select(cz_name, state, mobility, gini, social_capital_index,
            frac_foreign_born, frac_children_single_mothers, 
            school_expenditures_per_student, hs_dropout_rate_adj,
            urban, hh_income, racial_seg)
```


Take a few minutes to explore how some of the variables in the `cz_subset` data frame are associated with mobility. What is a relationship where you would expect a negative association? What is a relationship where you would expect a positive association? What is a relationship where you would expect no association?


### REPLACE THIS LINE WITH YOUR CODE



# Correlation Matrix

One way to save some time when looking at multiple correlations is to create a matrix with all the possible correlations in your dataframe. We can't use the first two variables in `cz_subset` since they are not numeric, but we can use variables 3 (`mobility`) through 12 (`racial_seg`). Let's get the matrix for correlations using all the rows and variables 3:12. In the chunk below, note the indexing with brackets. The blank space in front of the comma asserts we want to include all the rows, and the 3:12 after the comma asserts which columns we want to include: 

```{r create matrix}
matrix <- round(cor(cz_subset[ , 3:12], use = "complete.obs"),3) 
```

Let's review the matrix!

```{r review matrix}
matrix
```

To only see some of the matrix, use indexing!

```{r matrix with indexing}
matrix[1:4, 1:4]
```



# Introducing Multivariate Relationships - See Slides

Last week we identified individual points on a scatterplot to dig deeper into how two variables are associated. Another analytical tool is to examine whether the association differs at specific values of another variable.

Consider the correlation between racial segregation and income. Would you expect this correlation to be positive or negative? Strong or weak?

### REPLACE THIS LINE WITH YOUR CODE


Would this association be the same in urban and rural commuting zones?

### REPLACE THIS LINE WITH YOUR CODE


We can visualize the different associations by geographic type if we use different colored points for urban and rural commuting zones. Before we get there, let's make sure the `urban` variable is recognized as a factor variable, and change the labels from "0" and "1" to "Rural" and "Urban".

```{r make urban a factor variable}
cz_subset <- cz_subset %>%
     mutate(urban = factor(urban, 
                           levels=c("0", "1"), 
                           labels=c("Rural", "Urban")))
```
 
Now let's make the plot.

```{r urban rural plot}
urban_rural_plot <- ggplot(cz_subset, aes(x = hh_income, y = racial_seg, 
                                          color = urban, shape = urban))

urban_rural_plot + geom_point()
```


Extra Fancy: Remember we can use `facet_grid()` to display two plots in one figure. 

```{r urban rural facets}
urban_rural_facets <- ggplot(cz_subset, aes(x = hh_income, y = racial_seg))

urban_rural_facets + geom_point(aes(color = urban)) + 
     geom_smooth(method = lm, color = "black") + 
     facet_grid(.~urban) + 
     scale_color_manual(values = c("Red", "Blue")) + guides(color = FALSE) +
     labs(x = "Median Household Income", y = "Racial Segregation",
          title = "Rural-Urban Differences in Association Between \nMedian Household Income and Racial Segregation", 
          subtitle = "Data from Opportunity Insights")
```

Finally, let's create a three-way table showing how the mean of the racial segregation index varies at each quintile of median income between urban and rural commuting zones. Any thoughts on how to do this? What do we need?

### REPLACE THIS LINE WITH YOUR CODE 


In the table above we have long data but might prefer wide data. This would be a good time to use `spread()` rather than `gather()`. In the `spread()` line, we are saying we want to create a column for each level of the `urban` variable, and we want the values of those variables to be the values of `mean_racial_seg` for each quintile-urban combination.

```{r spread example}

cz_spread <- cz_quintiles %>%
     spread(urban, mean_racial_seg)

cz_spread
```

Pander will make this table prettier...

```{r pader the spread table}
pander(cz_spread)
```


# If there's time...

How could we visualize these three way relationships? Note that we'll want to use the long data for this rather than the wide data. Why?

One option for a plot is to use side by side *columns*...

```{r columns with dodge}
quintile_plot1 <- ggplot(cz_quintiles, aes(x = quintile, y = mean_racial_seg,
                                          fill = urban, 
                                          label = mean_racial_seg))

quintile_plot1 + geom_col(position = "dodge") + # For side by side columns
     geom_text(position = position_dodge(1), vjust = -.25) +
     guides(fill = FALSE)
```

Or side by side plots...

```{r columns with facets}
quintile_plot2 <- ggplot(cz_quintiles, aes(x = quintile, y = mean_racial_seg,
                                           fill = urban))

quintile_plot2 + geom_col() + 
     geom_text(aes(label = mean_racial_seg, vjust = -.25)) + 
     facet_grid(.~urban) + guides(fill = FALSE)
```


Want better colors? I like the [`viridis color palettes`](https://cran.r-project.org/web/packages/viridis/vignettes/intro-to-viridis.html).

```{r install and load viridis}
install.packages("viridis")
library(viridis)
```

```{r facet plot with viridis palette}
quintile_plot2 + geom_col() + 
     geom_text(aes(label = mean_racial_seg, vjust = -.25)) +
     facet_grid(.~urban) + guides(fill = FALSE) +
     scale_fill_viridis(discrete=TRUE) 

# Add discrete = TRUE since our fill variable is discrete (versus continuous)

```

