# Getting Started
We are still using the `gss_week3` dataset, and we will be using `ggplot2` and `dplyr` from `tidyverse`. Load the data and `tidyverse`, and remember to include the options to suppress the output, warning, and message from this chunk.

```{r load data and packages, echo = FALSE, warning = FALSE, message = FALSE}
gss_week3 <- read.csv("~/gss_week3.csv")
library(tidyverse)
```

On Tuesday, we ended by creating a new factor variable that broke the education distribution into categories. The first step was to define the variable name as an object and assign the values of `educ` as appropriate. Since we didn't save the data, we have to run that chunk again.

```{r set up education categories}
gss_week3$edcat[gss_week3$educ<12] <- "Less Than HS"
gss_week3$edcat[gss_week3$educ==12] <- "HS Diploma"
gss_week3$edcat[gss_week3$educ>12] <- "Some College"
gss_week3$edcat[gss_week3$educ==16] <- "College Degree"
gss_week3$edcat[gss_week3$educ>16] <- "Grad Prof Degree"
```

Next we made sure R knows that this is a factor variable, and ordered the categories by degree rather than alphabetically (the default):

```{r set up edcat as factor variable}
gss_week3 <- gss_week3 %>%
     mutate(edcat = factor(edcat,
                           levels = c("Less Than HS", 
                                      "HS Diploma", 
                                      "Some College", 
                                      "College Degree", 
                                      "Grad Prof Degree")))
```

We can use this new factor variable just like any other variable. For example, if you wanted to find the proportion of cases in each category, what would you do?

###REPLACE THIS LINE WITH YOUR CODE###


We have been working a lot with the center, spread, and shape of the distribution of age at first birth. As a refresher, can you make a density plot showing the distribution of that variable?

###REPLACE THIS LINE WITH YOUR CODE###

What possible explanations can you suggest for why age at first birth would vary by highest degree?

Let's get a sense of the distributions by overlaying density plots of `agekdbrn` for each level of `edcat`. We set everything up in ggplot2 the same way as before (but adjust the title as appropriate; note that `\n` forces a line break in the title). 

In the initial aesthetic mapping section, we add `color = edcat` to say that we want a separate plot in a different color for each education category:

```{r agekdbrn plot by edcat}
agekdbrn_edcat_plot <- ggplot(gss_week3, aes(x = agekdbrn, color = edcat))

agekdbrn_edcat_plot + geom_density() + 
  labs(x = "Age at First Birth", y = "Density", 
          title = "Distribution of Age at First Birth \n by Highest Degree", 
          subtitle = "GSS, 2000-2014")
```

There are lots of ways to improve this figure. You can check out the ggplot2 cheatsheet for some ideas. One option to consider is changing the title of the legend. You do so by adjusting the name of the scale that controls the colors:

```{r agekdbrn edcat plot legend title}
agekdbrn_edcat_plot <- ggplot(gss_week3, aes(x = agekdbrn, color = edcat))
agekdbrn_edcat_plot + geom_density() + 
  labs(x = "Age at First Birth", y = "Density", 
       title = "Distribution of Age at First Birth \n by Highest Degree", 
       subtitle = "GSS, 2000-2014") + 
  scale_color_discrete(name = "Highest Degree")
```

The overlapping plots are quite effective in this case. That's not always true. Another option would be to have a separate plot for each category. You can do that with faceting. In this example, we want to facet into rows with `.~edcat`. Note that with faceting you don't need a legend anymore. You can delete it using `guides(color=FALSE)`. 

```{r agekdbrn edcat plot with row facet}
agekdbrn_edcat_plot <- ggplot(gss_week3, aes(x = agekdbrn, color = edcat))
agekdbrn_edcat_plot + geom_density() + 
  labs(x = "Age at First Birth", y = "Density", 
       title = "Distribution of Age at First Birth by Highest Degree", 
       subtitle = "GSS, 2000-2014") +
facet_grid(edcat~.)  + guides(color=FALSE)
```

You can facet into columns by replacing `edcat~.` with `.~edcat`. Which would make more sense in this case?

```{r agekdbrn edcat plot with column facet}
agekdbrn_edcat_plot <- ggplot(gss_week3, aes(x = agekdbrn, color = edcat))
agekdbrn_edcat_plot + geom_density() + 
  labs(x = "Age at First Birth", y = "Density", 
       title = "Distribution of Age at First Birth by Highest Degree", 
       subtitle = "GSS, 2000-2014") +
facet_grid(.~edcat)  + guides(color=FALSE)
```


# Back to Basics
Let's take a break from graphics for a moment and think about tables and basic summaries again. If you wanted to find the proportion of respondents in each education category for a specific race/ethnicity category, what could you do?

###REPLACE THIS LINE WITH YOUR CODE###


Another option would be a cross-tabulation. We set up a cross-tabulation using the same syntax as the `table()` function, but now we'll have two variables separated by a comma. The values of our first variable will become the rows in the table, and the values of the second variable will become the columns. This ordering matters! Think of the row variable as your independent variable (the X) and the column as your dependent variable (the Y). If that's the setup, why would we want the values of `racehisp` to become the rows in this table and the values of `edcat` to become the columns?

```{r table racehisp edcat}
table(gss_week3$racehisp, gss_week3$edcat)
```

How do you think you make a cross-tabulation with proportions?

###REPLACE THIS LINE WITH YOUR CODE###


By default, these proportions are for the entire table. So the top left cell tells us that 3.2% of all the cases in this data frame are Black respondents who did not complete high school.

To get row proportions, add a comma and a 1 (since the first variable in our table command is the row variable) before closing the prop.table parentheses. 

```{r prop table row proportions}
round(prop.table(table(gss_week3$racehisp, gss_week3$edcat),1),3)
```

Each *row* in the above table sums to 1. To get column proportions, replace the 1 with a 2. In the table below, each *column* should sum to 1. In general, we'll use row proportions more than column proportions.

###REPLACE THIS LINE WITH YOUR CODE###


There are a couple other `dplyr` functions that help with finding values of one variable for all values of another variable. For example, we might want to find the mean age at first birth for each race/ethnicity category. To do so, we first need to name a new data frame where we will store our values. Then we group all the observations by the categorical variable. Then we create a variable for our new data frame that summarizes one variable for all the categories of our other variable. And we can connect these functions with pipes.

```{r group by and summarize means}
agekdbrn_racehisp <- gss_week3 %>% 
                     group_by(racehisp) %>% 
                     summarize(mean_agekdbrn = mean(agekdbrn))

agekdbrn_racehisp
```

This `summarize()` function works with a lot of the basic commands we have discussed (min, max, sd, var, etc.). Try finding the median age at first birth for each education category. You can still use the `agekdbrn_edcat` object:

###REPLACE THIS LINE WITH YOUR CODE###



# Introducing Plots With X and Y Variables

What can we do with this info? We might want to make a nice figure, or we might want to make a nice table. Let's do both.

First, the figure. What would your ideal image look like? (Don't scroll down...)









Let's imagine something like a bar for each education category that rises to that category's median age at first birth. For something like that, the values of the x axis will be the values of our `edcat` variable. And the values of the y axis will be the values of the `median_agekdbrn` variable from the data frame we just created. That's all ggplot2 needs to get us started:

```{r agekdbrn edcat plot}
med_agekdbrn_edcat_plot <- ggplot(agekdbrn_edcat, aes(x = edcat, y = median_agekdbrn))

med_agekdbrn_edcat_plot
```

When we wanted to count how many observations there were with each value of `agekdbrn` we used `geom_bar()`. But now we want each bar to represent actual values (not counts). To distinguish, ggplot2 uses `geom_col()` when we want a barplot's y axis to display values rather than counts: 

```{r agekdbrn edcat plot with columns}
med_agekdbrn_edcat_plot + geom_col()
```

Adding a text label displaying the height of each column would be a good idea:

```{r agekdbrn edcat plot with labels}
med_agekdbrn_edcat_plot + geom_col() + geom_text(aes(label = median_agekdbrn))
```

And we'll adjust the labels outward from the columns so they are easier to read:

```{r agekdbrn edcat plot with outward labels}
med_agekdbrn_edcat_plot + geom_col() + geom_text(aes(label = median_agekdbrn,
                                                     vjust="outward"))
```

Try fixing the axes and give your plot a title and subtitle:

###REPLACE THIS LINE WITH YOUR CODE###


Graphics with ggplot2 get really fun when we start adjusting colors. The `RColorBrewer` package offers many colors beyond the defaults. Install the package and load it. (Remember that next time you use it, you do not have to install it again, just laod it.)

```{r load RColorBrewer}
# Install using the "Packages" tab in the bottom right pane
library(RColorBrewer)
```

To see all the colors options available through the `RColorBrewer` package, use the following code:

```{r see color options}
display.brewer.all()
```


To adjust the colors in the plot using one of these palettes, use the `scale_fill_brewer()` option:

```{r change color palette}
plot <- ggplot(agekdbrn_edcat, aes(x = edcat, y = median_agekdbrn, fill = edcat))

plot + geom_col() + scale_fill_brewer(palette="Spectral")
```

That figure looks great now! Can we make a table look as nice?

\pagebreak

# Improving Tables with Pander

There are different options for improving the format of tables. One of the easiest options is to use the `pander` package. Install it and load it.

```{r load pander}
# Install using the "Packages" tab in the bottom right pane
library(pander)
```

The `pander` package manipulates the *pandoc rendering* of R output so it is more Markdown friendly. The most straightforward way to use this package is to wrap your code in the `pander()` function.

```{r pander table}
pander(agekdbrn_edcat)
```

Don't worry: it looks weird in the notebook because it is still in Markdown format. Also, you may want to change the column names from variable names to real names. You can do that with the `colnames()` function. (There is also a `rownames()` function if you need it in the future.):

```{r change table's column names}
colnames(agekdbrn_edcat) <- c("Highest Degree", "Median Age at First Birth")
pander(agekdbrn_edcat)
```

Okay, it still looks weird. But knit a PDF and check out the last page...
