# Setting Up
This week we will continue using the 2000-2014 waves of the General Social Survey. We are still focused on the `agekdbrn` variable but will start looking at how its distribution varies by education and race/ethnicity. Start by loading tidyverse and the data.

```{r load packages and data, echo = FALSE, message = FALSE}
library(tidyverse)
gss_week3 <- read.csv()
```


# Variance and Standard Deviation

To find the variance, use `var()`. To find the standard deviation, use `sd()`.

```{r variance and sd of agekdbrn}
var(gss_week3$agekdbrn)
sd(gss_week3$agekdbrn)
```

Would you expect more or less variation in the education distribution?

### REPLACE THIS LINE WITH YOUR CODE

```{r variance and sd of educ}
var(gss_week3$educ)
sd(gss_week3$educ)
```


# Introducing ggplot2
Let's look at some visualizations of the distribution of age at first birth. There are two main types of visualizations you can make with R: base graphics and ggplot2 graphics. We are going to focus on ggplot2, one of the packages (like dplyr) that loads with the tidyverse.

The "grammar of graphics" that ggplot is built on creates graphics layer by layer, mapping aesthetics and options on top of other layers. Let's see how this works in a very basic density plot.

The first step is to name our figure as an object called `agekdbrn_plot`. The right side of the following chunk says that we will be using the ggplot functions with the `gss_week3` data frame, and the `aes` parentheses is where we say what we want on the x and y axes. For this initial plot, we will only have a variable on the x axis.

```{r save plot as object}
agekdbrn_plot <- ggplot(gss_week3, aes(x = agekdbrn))
```

To get a sense of how plots are built layer by layer, let's take a look at what is now connected to our object:

```{r view plot background}
agekdbrn_plot
```

So R knows the range of our variable and has used that info to set the backdrop for our plot. The next step is to add a layer on top of this. We use our existing object name and add the geometic feature we want. In this example, we will look at a density plot:

```{r add density plot layer}
agekdbrn_plot + geom_density()
```

What is this distribution telling us? How would you define the shape and the skew?

There are many ways to improve this plot. We'll start simply by changing the axis labels and adding a title. If you are making all of these changes, the easiest thing to do is to use the `labs()` function:

```{r add labels}
agekdbrn_plot + geom_density() + 
  labs (x = "Age at First Birth", y = "Density",
        title = "Distribution of Age at First Birth",
        subtitle = "GSS, 2000-2014")
```

Two other visualizations that are good for looking at distributions are bar plots and histograms. These use `geom_bar` and `geom_histogram` respectively. Try them out by editing the last code chunk we ran to replace `geom_density` with `geom_bar` or `geom_histogram`:

### REPLACE THIS LINE WITH YOUR CODE

The warning message about binwidth with the histogram is asking you to choose how many buckets you want the data grouped by. See what happens as you increase and decrease the number of bins:

```{r agekdbrn histogram bins}
agekdbrn_plot + geom_histogram(bins = 40) + 
  labs (x = "Age at First Birth", y = "Density",
        title = "Distribution of Age at First Birth",
        subtitle = "GSS, 2000-2014")
```


# Creating Categories

We already know how to use indexing to get the values of one variable for observations that have specific values of another variable. That would be a good start for finding the average age at first birth for different ranges of the education distribution. But it would be more helpful if we could collapse the values of the education variable into categories. 

We'll do this by creating a new object in our data frame (called `edcat` in this example), and we'll create new values of this variable according to indexed values of the existing `educ` variable. For example, in this next line we are saying that the new `edcat` variable should take the value "Less Than HS" if the values of the existing `educ` variable are less than 12:


```{r creating categories}
gss_week3$edcat[gss_week3$educ<12] <- "Less Than HS"
```

We want to create the following education categories: "HS Diploma", "Some College", "College Degree", and "Grad Prof Degree". How should we cut the values of the `educ` variable to make these categories?

###REPLACE THIS LINE WITH YOUR CODE

The next step is to tell R we will be using this new variable as a factor variable. We can do this using `mutate()`. We also need to assert the order of the categories, since by default the categories will be listed alphabetically. To assert the oder, we will add the `levels = c()` option. 

```{r factor variable with levels}
gss_week3 <- gss_week3 %>%
     mutate(edcat = factor(edcat,
          levels = c("Less Than HS",
                     "HS Diploma",
                     "Some College",
                     "College Degree",
                     "Grad Prof Degree")))
```

Now we are at the point where we could summarize each category by name rather than by range of years:

```{r summary by category}
summary(gss_week3$agekdbrn[gss_week3$edcat=="College Degree"])
```

What is the standard deviation of `agekdbrn` for respondents with a graduate or professional degree? What is the standard deviation of `agekdbrn` for respondents who completed less than high school?

###REPLACE THIS LINE WITH YOUR CODE

On Thursday, we will visually compare the distributions of `agekdbrn` for each education category.

