## Setting Up

We''ll use the `midd_survey` data for this example. Load it and the usual packages.

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(pander)
midd_survey <- read.csv("midd_survey.csv")
```


## Formatting Regression Tables With Stargazer

Summaries of linear regression models have a lot of information and can be messy in knitted files. One option to clean them up is to use the `stargazer` package. The `stargazer` package is highly customizable. We'll go over just a few options today. For more details about the package, click [here](http://jakeruss.com/cheatsheets/stargazer.html). 

To see how stargazer compares to the output we are used to, use the `midd_survey` data to regress `gpa` on `siblings`, save the model as `model1`, and look at the summary:

```{r}
model1 <- lm(gpa ~ siblings, data = midd_survey)
summary(model1)
```

Now that you are comfortable interpreting summary outputs, you should continue using the traditional summary outputs to get a sense of your models. But replace `summary()` with `stargazer()` when you want to knit your file. 

```{r}
# INSTALL STARGAZER FIRST IN THE BOTTOM RIGHT PANE!
library(stargazer)
stargazer(model1)
```

There are a few other things to add to make sure the file knits properly:

- `results = "asis"` is the most important thing. It leaves the output in the format that knitr needs to render it. (Add this option to the open fence line; add everything else to the stargazer function line.)
- `header = FALSE` deletes the contact information for the package author
- `type = "latex"` converts the output into the Latex language that pandoc needs to knit to a PDF. Latex is the default so you can leave it off if you want. To change to web out put you would want to change `type = "latex"` to `type = "html"`.


```{r, results = 'asis', echo = FALSE}
stargazer(model1, header = FALSE, type = "latex")
```

Knit the file to see the formatted regression table. This is better than dumping raw R output into a report, but we can get better. First, let's add a model with a control variable and a model with an interaction.

```{r, include = FALSE}
model2 <- lm(gpa ~ siblings + gender, data = midd_survey)
model3 <- lm(gpa ~ siblings * gender, data = midd_survey)
```

Here are some more changes for the `stargazer()` function:

- `style = "ajs"` formats the table following publication guidelines for the *American Journal of Sociology*. Replace "ajs" with "asr" to use the formatting style for the *American Sociological Review*. These are the two most common formatting styles you will see for sociology papers. I prefer ajs style for written reports (because it includes standard errors), and asr style for slides (because it is more condensed).
- `title` creates a title for the table
- `covariate.labels = c()` is a list of names for the independent and control variables in your model. You need to label every variable or this won't work.
- `dep.var.labels = ` creates a title above the dependent variable model numbers
- `keep.stat = c("n")` reports the number of observations but excludes all the other model information

```{r, results = "asis", echo = FALSE, warning = FALSE}
stargazer(model1, model2, model3,
          header=FALSE, type = "latex", style = "ajs",
  title = "A Better Title",
  covariate.labels = c("Number of Siblings", "Gender = Other", "Gender = Woman",
                       "Siblings X Gender = Other", "Siblings X Gender = Woman"),
  dep.var.labels = "GPA",
  keep.stat = c("n"))
```