## Getting Started

To begin, copy all the text on this screen. Then log in to https://rstudio.middlebury.edu, and in the "File" menu select "New File" then "R Markdown". Add a title, include your name in the "Author" box, and click Ok. A default markdown file will open. Delete all the text in the default file and paste in the text you copied to your clipboard. Make sure that "R Markdown" is selected in the file type drop down menu in the bottom right corner of this pane. Save the file (in the R menu, not your browser menu!) in your "Home" folder on the R Studio server. 


## Once Your File Is Saved

We are using a dataset containing information pulled from nearly 60,000 OK Cupid profiles. The online dating profiles were created by users who were active online in June 2012 and who lived within 25 miles of San Francisco, CA at that time.

This dataset and the R package to go with it was made publicly available by Albert Kim in 2015.

Let's load the packages and the data and clean up a couple variables:

```{r load packages and data, warning=FALSE, error=FALSE, message=FALSE}
#install.packages("okcupiddata")
#install.packages("tm")
#install.packages("wordcloud")
library(okcupiddata)
suppressMessages(library(dplyr))
library(ggplot2)
suppressMessages(library(wordcloud))
suppressMessages(library(tm))
library(stringr)

data(profiles)

profiles <- profiles %>% filter(height>=55 & height <=80, age<=75) %>%
     mutate(ethnicity = ifelse(grepl(",",ethnicity), "multiple", as.character(ethnicity))) %>%
     mutate(ethnicity = ifelse(ethnicity == "pacific islander", "asian", ethnicity)) %>%
     mutate(ethnicity = ifelse(ethnicity == "native american", "other", ethnicity)) %>%
     mutate(ethnicity = ifelse(ethnicity == "middle eastern", "other", ethnicity))
```


## Are online daters representative?

Whenever data about a *sample* is used to make claims about a *popluation*, it is important to know how representative the sample is of the population. One way to do this is to test whether sample characteristics are significantly different from measures known about the population. 

As an example, let's look at the distribution of height in the OKC profiles. Why might height be a good way to compare our sample to the population? Why might height be less helpful?

```{r summarizing height}
summary(profiles$height)
```

The distribution of height in the population varies by sex (among other characteristics). The average height for adult females in the United States is 63.8 inches and the average height for adult males in the United States is 69.7 inches. How does the distribution of height for female users in our sample compare to the national average?

```{r summarizing height for females}
summary(profiles$height[profiles$sex=="f"])
```

So our sample is slightly taller than the national average. Why might that be?

The next step is to know if the difference in height between the sample and the population is statistically significant. One common tool is a t-test to test whether the difference in means between two groups is outside the range of a confidence interval. 

```{r ttest of height for females}
t.test(profiles$height[profiles$sex=="f"], mu = 63.8)
```

There are a few key pieces of information here that show us that females in our sample are *significantly* taller than females in the population. In formal language, these results suggest we can *reject the null hypothesis that the difference in means is zero*. 

First, note the test statistic (t-value) of 73.742. That tells us the difference we observe is very far from what we know the population mean is. Once the t-stat is more extreme than +/- 1.96, we are comfortable rejecting the null hypothesis at the 95% confidence level.

Second, note the p-value of <2.2e-16. That tells us there is a very small probability of randomly drawing another sample that would have a mean height that is taller than our sample's mean height. Once the p-value is less than .05, we are comfortable rejecting the null hypothesis at the 95% confidence level.

Third, note the 95 percent confidence interval of [65.07386, 65.14343]. That tells us that our dataset would predict that if we took repeated samples of the population that are the same size as our sample we would expect the mean heights from all those samples to fall within that confidence interval 95% of the time. Since the known population mean of 63.8 is not within that range, we are comfortable saying the sample mean and the population mean are significantly different.

What about male users? Would you expect their mean height to be significantly different from the population mean of 69.7? Find out below...


```{r height for males}
summary(profiles$height[profiles$sex=="m"])

t.test(profiles$height[profiles$sex=="m"], mu = 69.7)
```


##Describing the Sample

Let's use simple frequency tables and proportions to learn more about our sample.

We already know the heights of male and female users differ. What about their ages? Summarize the distribution of age by sex in this sample.

```{r age summary by sex}
summary(profiles$age[profiles$sex=="f"])
summary(profiles$age[profiles$sex=="m"])
```

They actually look very similar! Is the difference in means significant? To find out, we can test the observed difference against the null hypothesis that the true difference is zero.

```{r testing sex differences in mean age}
t.test(profiles$age[profiles$sex=="f"], profiles$age[profiles$sex=="m"])
```









The mean is only one summary statistic. And in some cases, only looking at the mean will obscure more interesting insights. Whenever possible, it is preferable to look at the full distribution of values.

How would you expect the age distribution of OKC users to vary by sex? We know the average age is very similar...but would you expect older or younger users to also be spread evenly by sex? 

Let's see how the shape of the age distribution varies by sex:

```{r plot of age distribution by sex}
age_sex_plot <- ggplot(profiles, aes(x = age, color = sex))
age_sex_plot + geom_density()
```


Let's move to self-reported sexual orientation. What proportion of users would you expect to list their sexual orientation as bisexual, gay, or straight?

We can look at a frequency table to start:

```{r frequency table of orientation}
table(profiles$orientation)
```

And then wrap that table in the `prop.table()` function to get proportions. We'll also round the proportions to three decimal places:

```{r proportion table of orientation}
round(prop.table(table(profiles$orientation)),3)
```

That is the distribution across all users. How might it vary by sex?

```{r orientation proportions by sex}
round(prop.table(table(profiles$sex, profiles$orientation)),3)
```

Are these the proportions we want? What would proportions by row or by column tell us? Add a 1 for row proportions (because the variable in the rows is the first variable in our command). Add a 2 for column proportions (because the variable in the rows in the second variable in our command).

```{r row and column orientation proportions by sex}
round(prop.table(table(profiles$sex, profiles$orientation),1),3)
```

Let's turn to the `ethnicity` variable. How does sexual orientation differ by ethnicity in this sample? We'll start with a simple proportion table. Remember to pay attention to whether you want row or column proportions!

```{r ethnicty and orientation proportions}

```


As one more step, let's combine sex, ethnicity, and orientation. Once we get to three variables it is easier to make a flat table using `ftable()` rather than a regular table.

```{r sex and ethnicity and orientation proportions}
round(prop.table(ftable(profiles$sex, profiles$ethnicity, profiles$orientation),1),3)
```

Finally, let's see how to plot relationships among three variables. In this next chunk we want to see how the age distribution by sex that we saw before varies by sexual orientation. How would you interpret these distributions?

```{r age orientation plot}
age_orientation_plot <- ggplot(profiles, aes(x = age, color = orientation))
age_orientation_plot + geom_density() + facet_grid(sex~.)
```


## Analyzing Text Responses

Now we'll examine trends in excerpts from one of the essays in the OKC profiles. Let's clean up some of the variables so they are ready to use:

```{r word setup}
essays <- select(profiles, starts_with("essay"))
essays <- apply(essays, MARGIN=1, FUN=paste, collapse=" ") 
essays <- str_replace_all(essays, "\n", " ")
essays <- str_replace_all(essays, "<br />", " ")

profiles <- profiles %>% mutate(kids = ifelse(grepl("has", offspring), "has kids", "no kids"))
```

We'll start by creating a list of the words that appear most frequently in profiles written by female users, and then look at the top 25. What is a word you expect we'll find in the top 25?

```{r female words}
female.words <- subset(essays, profiles$sex=="f") %>% str_split(" ") %>%
     unlist() %>%
     table() %>%
     sort(decreasing=TRUE) %>%
     names()

female.words[1:25]
```

So those words are not very exciting. But this is a step toward something more exciting. Next get a list of the words that appear most frequently in profiles written by male users.

```{r male words}
male.words <- subset(essays, profiles$sex=="m") %>% str_split(" ") %>%
     unlist() %>%
     table() %>%
     sort(decreasing=TRUE) %>%
     names()

male.words[1:25]
```

Now the fun part: let's get a list of the 25 words that most frequently appear in female users' profiles that *do not* appear in male users' profiles:

```{r female not male words}
female_male_word_differences <- setdiff(female.words[1:500], male.words[1:500])
female_male_word_differences[1:25]
```

Find the list of 25 words that most frequently appear in male users' profiles that *do not* appear in female users' profiles:

```{r male not female words}


```

Lastly, let's make a word cloud of the 100 most common female words. You will probably need to make this window a lot bigger to be able to see the result:

```{r female wordcloud}
wordcloud(profiles$essay0[profiles$sex=="f"], 
max.words = 100, scale=c(5,.5),random.order = TRUE, colors=brewer.pal(8, "Dark2"))
```



## Next Steps?

What else would you want to consider with word differences or word clouds? There are several other variables in the dataset that could be interesting to explore...

ethnicity

age

orientation

kids

diet

body_type

religion

To see the categories or levels for them, wrap them in a table:

```{r}
table(profiles$kids)
```
